Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motion Characteristic to Vertical Axis Turbine

### 2016-02-29 abstract.tex
Use my full name
Wrap abstract text

\title{Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motion Characteristic to Vertical Axis Turbine} % \\ \jpcs}
\author{Victor Mendoza$^{1}$, Pete Bachant$^{2}$, Martin Wosnik$^{2}$ and Anders Goude$^{1}$ }
\address{$^{1}$ Department of Engineering Sciences, Division of Electricity, Uppsala University, \\Uppsala 751 21, Sweden}
\address{$^{2}$ Center for Ocean Renewable Energy, University of New Hampshire, 24 Colovos Rd.,\\ Durham, NH 03824, USA}
\ead{victor.mendoza@angstrom.uu.se}

\begin{abstract}
Vertical Axis Turbines (VAWT) have been used for extracting of energy either from sea or wind. A simpler design, low cost of maintenance and the capability of working independent from the direction of the flow are some of the most important advantages over conventional Horizontal Axis Turbines (HAWTs). Complex and unsteady fluid dynamics are not easy to model for simulation tools. One of the most relevant phenomena is the dynamic stall which is inherently to unsteady conditions, this work is focused in its study. 
Dynamic stall is usually used as a passive control for VAWT operating conditions, hence the importance of a good understanding. In this study, \textbf{a coupled model is implemented with the open-source CFD toolbox OpenFOAM for solving the governing equations; an actuator line model is used for solving velocity flow field and a dynamic stall model for force interactions.} Force coefficients obtained from the model are validated with experimental data of pithing airfoil in similar operating conditions as H-rotor type.   
\end{abstract}

>>>>

\title{Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motion Characteristic to Vertical Axis Turbine} % \\ \jpcs}
\author{Victor Mendoza$^{1}$, Peter Bachant$^{2}$, Martin Wosnik$^{2}$ and Anders Goude$^{1}$ }
\address{$^{1}$ Department of Engineering Sciences, Division of Electricity, Uppsala University, \\Uppsala 751 21, Sweden}
\address{$^{2}$ Center for Ocean Renewable Energy, University of New Hampshire, 24 Colovos Rd.,\\ Durham, NH 03824, USA}
\ead{victor.mendoza@angstrom.uu.se}

\begin{abstract}

    Vertical Axis Turbines (VAWT) have been used for extracting of energy either
    from sea or wind. A simpler design, low cost of maintenance and the
    capability of working independent from the direction of the flow are some of
    the most important advantages over conventional Horizontal Axis Turbines
    (HAWTs). Complex and unsteady fluid dynamics are not easy to model for
    simulation tools. One of the most relevant phenomena is the dynamic stall
    which is inherently to unsteady conditions, this work is focused in its
    study. Dynamic stall is usually used as a passive control for VAWT operating
    conditions, hence the importance of a good understanding. In this study,
    \textbf{a coupled model is implemented with the open-source CFD toolbox
    OpenFOAM for solving the governing equations; an actuator line model is used
    for solving velocity flow field and a dynamic stall model for force
    interactions.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as
    H-rotor type.

\end{abstract}
